```{r, include = FALSE, cache = FALSE}
library(knitr)
# opts_knit$set(out.format = "latex")
knit_theme$set(knit_theme$get("greyscale0"))

# options(replace.assign=FALSE,width=50)

# opts_chunk$set(fig.path='figure/graphics-', 
#                cache.path='cache/graphics-', 
#                dev='pdf', fig.width=5, fig.height=5, 
#               cache=FALSE)
# knit_hooks$set(crop=hook_pdfcrop)

# figure referencing hack
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- paste0("Figure ",i)
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})

# usage 
# chunk options fig.cap = fig$cap(<label>, <caption>)
# reference `r fig$ref(<label>)`
```

## A suggestion

This is just a suggestion to try and help get you into "good habits" early. If I was sitting to take this practical now, I would start a new R script file. That way all of the work that you have done associated with today's course is in one place, and the code for each of the practicals is separate from one another. This might feel a bit tedious right now, but as the amount of code you write and number of projects you take part in increases it will pay off to have a structured workflow. 


# Data frames 

For this set of questions we will use the movies data from the IMDB database. This data is contained in a package called `ggplot2movies`. To load the data into your current session you can use

```{r}
data(movies, package = "ggplot2movies")
```

(1)  Use the `head()` function to inspect the top of the data, this can help give you a feel for what the data looks like and what variables are contained within the data?

    ```{r}
    head(movies)
    ```

(1) How many films and how many variables are in this data set?
    
    ```{r}
    dim(movies)
    ```

(1) Recall that if I want to pull out a single column, or variable, from a data frame we can use `$`. To extract the titles from this data set we write `movies$title`.^[If you can't remember what the names of the columns are, you can use `colnames(movies)` to find out.] Using `mean()` and `median()` calculate these summary statistics for the film lengths

    ```{r}
    mean(movies$length)
    median(movies$length)
    ```
    
(1) How many Action films are there in the data? (There is a 1 in the action column if it is an action film)

    ```{r}
    table(movies$Action)
    ```

(1) You can use `table()` on more than one column at once by just adding extra columns as arguments. For example `table(movies$mpaa, movies$Action)` would tell me how many action and non action films there are for each of the mpaa ratings. How many romantic comedy films are there? i.e a 1 in the Romance column and a 1 in the Comedy column?

    ```{r}
    table(movies$Romance, movies$Comedy)
    ```

(1) When the values are similar in each variable in the table, it can be difficult to tell which dimension (rows or columns) corresponds to which variable. To aid this there is an argument to the `table()` function, `dnn` which allows us to provide a vector of names. Try 

    ```{r}
    table(movies$Romance, movies$Comedy, dnn = c("Romance","Comedy"))
    ```


(1) What year is the oldest film in the data set from?

    ```{r}
    min(movies$year)
    ```

(1) How long is the longest film?

    ```{r}
    max(movies$length)
    ```

(1) What is the standard deviation of the film ratings?

    ```{r}
    sd(movies$rating)
    ```

(1) How many people cast their vote on a film on average?

    ```{r}
    mean(movies$votes)
    ```

(1) We have seen that `sort()` can be used to arrange values in a vector into ascending or descending order. With data frames we are often interested in being able to order the data set based on the order of one of the variables. For example, rearranging the movies data by year of release. To accomplish this we can use `order()`. `r knitr::inline_expr("movies[order(movies$year),]")` would acheive this example. Give it a try.

The following question is a little harder:

(1) What is the average length of the 50 best (highest rating) films? How does this compare to the average of all films.

    ```{r}
    # rearrange the data
    movies_sorted = movies[order(movies$rating, decreasing = TRUE),]
    # take the best 50
    best = movies_sorted[1:50,]
    # calculate average length
    mean(best$length)
    # compared to overall average length
    mean(movies$length)
    # perhaps surprisingly 30 mins shorter average for the best films
    ```


# Solutions

Solutions to the practical questions are contained within the package

```{r, eval= FALSE, echo = TRUE}
vignette("solutions2", package = "jrIntroduction")
```
