```{r, include = FALSE, cache = FALSE}
library(knitr)
# opts_knit$set(out.format = "latex")
knit_theme$set(knit_theme$get("greyscale0"))

# options(replace.assign=FALSE,width=50)

# opts_chunk$set(fig.path='figure/graphics-', 
#                cache.path='cache/graphics-', 
#                dev='pdf', fig.width=5, fig.height=5, 
#               cache=FALSE)
# knit_hooks$set(crop=hook_pdfcrop)
suppressPackageStartupMessages(library(dplyr))

# figure referencing hack
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- paste0("Figure ",i)
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})

# usage 
# chunk options fig.cap = fig$cap(<label>, <caption>)
# reference `r fig$ref(<label>)`
```

# Data manipulation 

This section will hopefully help you get more comfortable with some of the `dplyr` functionality for "wrangling" your data. We will do some data wrangling and the use that to create some plots. Make sure you load in the `dplyr` package and the `movies` data set.

```{r, echo = 2:3, eval = 2:3}
suppressPackageStartupMessages(library(dplyr))
library(dplyr)
data("movies",package = "ggplot2movies")
```


(1) We want to look at how film budgets have changed over time for both the Comedy and non Comedy films. To start with, since there are missing values `NA` for films where we didn't know the budget, there could be complications for calculating averages and plots. We can use the `is.na()` function to find such values. Use the `filter()` function to remove them, try using the `%>%` operator notation if you like, `movies %>% filter(!is.na(...))`

    ```{r, eval = FALSE}
    movies %>%
      filter(!is.na(budget))
    ```


(1) We want to look at comedy and non comedy films in each year, this is some sort of grouping structure which suggests use of the `group_by()` function. Create this grouping structure on the filtered movies data

    ```{r, eval = FALSE}
    movies %>% 
      filter(!is.na(budget)) %>%
      group_by(year,Comedy)
    ```

(1) Use `summarise()` to calculate the average budget in each year for both comedy and non comedy films. When you do name the new column of averages `b` and store the final output data as a variable called `plotdat`

    ```{r}
    plotdat = movies %>% 
      filter(!is.na(budget)) %>%
      group_by(year,Comedy) %>%
      summarise(b = mean(budget))
    ```

<!-- (1) We could now create a line plot for the non comedy movies using^[Here we are using the formula notation for plots, refer back to the plotting section of the notes if you are unsure.] -->

<!-- ```{r, echo=FALSE, eval = TRUE} -->
<!-- non_comedy = plotdat %>%  -->
<!--       filter(Comedy == 0) -->
<!-- ``` -->


<!--     ```{r, echo = TRUE, eval = FALSE} -->
<!--     non_comedy = plotdat %>%  -->
<!--       filter(Comedy == 0) -->
<!--     plot(b~year, data = non_comedy, type = "l",  -->
<!--          ylab = "Average Budget") -->
<!--     ``` -->

<!-- (1) Use `lines()` to add the corresponding line for the comedy films to be able to compare how the average budgets for each film type compares over time. -->

<!-- ```{r, eval = TRUE, echo = FALSE} -->
<!-- comedy = plotdat %>%  -->
<!--       filter(Comedy == 1) -->
<!-- ``` -->

<!--     ```{r, eval = FALSE} -->
<!--     comedy = plotdat %>%  -->
<!--       filter(Comedy == 1) -->
<!--     lines(b~year, data = comedy, col = 2) -->
<!--     ``` -->

<!-- (1) Similarly create a data set which calculates the average rating for different film lengths for both comedy and non comedy films -->
<!-- called `films` -->

<!--     ```{r} -->
<!--     films = movies %>% -->
<!--       group_by(length,Comedy) %>% -->
<!--       summarise(r = mean(rating)) -->
<!--     ``` -->

<!-- (1) We could then plot the average rating against length for the non comedy films using the code below. In my data the average ratings column is called `r` -->

<!-- ```{r, eval=TRUE, echo = FALSE} -->
<!-- non_films = films %>% -->
<!--            filter(Comedy == 0) -->
<!-- ``` -->


<!--     ```{r, echo = TRUE, eval = FALSE} -->
<!--     non_films = films %>% -->
<!--            filter(Comedy == 0) -->
<!--     plot(r ~ length, data = non_films, pch = 19, -->
<!--          ylab = "Average Rating") -->
<!--     ``` -->

<!-- (1) Add the points for the Comedy films in red using the `points()` function -->

<!-- ```{r, echo = FALSE} -->
<!-- com_films = films %>%  -->
<!--       filter(Comedy == 1) -->
<!-- ``` -->


<!--     ```{r, eval = FALSE} -->
<!--     com_films = films %>%  -->
<!--       filter(Comedy == 1) -->
<!--     points(r~length, data = com_films, pch = 19, col = 2) -->
<!--     ``` -->
